---
title: "R Tutorial #3: Cryto/Stock Return Data, Calculating Returns & Uncorrelated Clusters"
output: html_notebook
---

# Packages that will be used
```{r}
library(tidyverse) # collection of packages for data wrangling & data science
library(crypto2) # an easy way to pull crypto currency data
library(lubridate) # working with date/times
library(rvest) # web scraping
library(stats) # statistics package
library(magrittr) # operators that reduce development time 
library(quantmod) # package useful for quantitative trading & research
library(tidyquant) # a wrapper with convenient functions 
library(dendextend) # a package for creating visually appealing dendrograms
library(PortfolioAnalytics) # a package to quickly evaluate portfolio performance
library(janitor) #used in the tutorial for the "clean_names()" function
```


#SECTION 1 - GET A LIST OF ALL CRYPTO TICKERS, AND THEIR PRICE HISTORY
```{r}

crypto_list() #lists all cryptocurrencies

crypto_currency <- crypto_history(limit = 1) #pulls price history for one currency (most likely BTC)

crypto_currency %>% 
  mutate(timestamp = as.Date(as.character(timestamp))) %>% 
  ggplot()+
  geom_line(aes(timestamp, close)) #visualize the price history

crypto_currency$name %>% unique() #check how many unique cryptos are in your dataset

top_100_by_marketcap <- 
crypto_list() %>% #list all cryptos agaon
  arrange(rank) %>% #arrange the based on market cap rank
  slice(1:100) #select rows 1 to 100

top_100_crypto_prices <-
crypto_history(top_100_by_marketcap) %>% #get data for all 100 coins
  mutate(timestamp = as.Date(as.character(timestamp))) #fix the timestamp into a date object

```

#SECTION 2 - CALCULATE CRYPTOCURRENCY RETURNS
```{r}

crypto_daily_returns <- 
top_100_crypto_prices %>% 
  arrange(symbol, timestamp) %>% #make sure to arrange the data first so the lag calculations aren't erroneous
  group_by(symbol) %>%  
  mutate(daily_return = close/lag(close, 1)-1) %>% #calculate the return in prices
  select(timestamp, name, symbol, daily_return) #select a subset of the columns - not 100% neccecary

crypto_daily_returns #view the final results

```

#SECTION 3 - WORKING WITH HIERARCHICAL CLUSTERING ALGORITHM
```{r}

hc <- 
crypto_daily_returns %>% 
  pivot_wider(id_cols = timestamp, names_from = name, values_from = daily_return) %>% #make the data wide, instead of long
  select(-timestamp) %>% #remove the timestamp - we want to exclude these from the calculation
  cor(use = 'complete.obs') %>% #correlation matrix 
  abs() %>% #absolute value
  dist() %>% #distance matrix
  hclust() #hierarchical clustering

hc %>% 
  as.dendrogram() %>% #convert clustering object into dendrogram
  plot() #view the results

number_clusters <- 3 #how many clusters do you want to select

#use the ggdend function to make the chart workable in ggplot
hc %>% 
  as.dendrogram() %>% 
  color_branches(k = number_clusters) %>% 
  color_labels(k = number_clusters) %>% 
  set('labels_cex', 0.5) %>% 
  # plot()
  as.ggdend() %>% 
  ggplot() +
  labs(title = 'Dendrogram of the top 100 Cryptocurrencies by market cap')

```

#SECTION 4: CLASSIFY EACH CRYPTO INTO A CORRESPONDING CLUSTER
```{r}

#extract the cluster values for each cryptocurrency
cutree(hc, k = number_clusters) %>% 
  as.data.frame() %>% 
  rename(cluster = 1) %>% 
  mutate(token_name = rownames(.)) %>% 
  filter(cluster == 3)

```

#SECTION 5: GET ALL PRICE HISTORY FOR S&P500 COMPANIES
```{r}

#example of the tidyquant function, visualizing apple's stock price
tidyquant::tq_get('AAPL') %>% 
  ggplot() +
  geom_line(aes(date, adjusted))

stock_table_url <- 'https://en.wikipedia.org/wiki/List_of_S%26P_500_companies' #url that has all stock tickers in the S&P500

#scrape the data and convert it into a table
all_historical_spx_ticker <- 
stock_table_url %>% 
  read_html() %>% 
  html_nodes(css = 'table') %>% 
  extract() %>% 
  html_table(fill = T)

# all_historical_spx_ticker[2] #view the delisted / changed ticker list

# assign list of tickers to a dataframe
current_spx_tickers <- 
 all_historical_spx_ticker[[1]] %>% 
   janitor::clean_names() 

pull_all_data <- . %>% print() %>% tq_get() %>% as.data.frame() #function that prints the ticker, then returns the data

tq_get('BRK.B') #example of BRK.B - this won't work, so on line 137 below we adjust the ticker string

'MMM' %>% pull_all_data() #example of the function above in action

all_spx_prices <- #provides price history for all tickers in the S&P500
current_spx_tickers %>% 
  mutate(symbol = stringr::str_replace_all(string = symbol, pattern = '[.]', replacement = '-')) %>% 
  mutate(data = map(symbol, pull_all_data)) %>% 
  select(-symbol) %>%
  unnest_legacy()

# all_spx_prices <- read_csv('spx_historical_prices.csv')

# all_spx_prices %>% 
#   select(-symbol) %>% 
#   unnest_legacy() %>% #pull(symbol) %>% unique() 

```


#SECTION 6 PASS THEM THROUGH THE HCLUST ALGO
```{r}

#this chunk is a repeat of the crypto returns, but for stocks. view the dendrogram to see the different clusters

wide_stock_returns <- 
all_spx_prices %>% 
  select(date, security,symbol, adjusted) %>% 
  arrange(symbol, date) %>% 
  group_by(symbol) %>%
  mutate(daily_return = adjusted/lag(adjusted)-1) %>% 
  na.omit() %>% 
  pivot_wider(id_cols = date, names_from = security, values_from = daily_return) 

hc <- 
wide_stock_returns %>% 
  select(-date) %>% 
  cor(use = 'complete.obs') %>% 
  dist() %>% 
  hclust()


hc %>% 
  as.dendrogram() %>% 
  color_branches(k = 6) %>%
  color_labels(k = 6) %>%
  set('labels_cex', 0.5) %>% 
  # plot()
  as.ggdend() %>% 
  ggplot() +
  labs(title = 'Dendrogram of the S&P500 Constituents')


```

#SECTION 7 NAIVE PORTFOLIO CONSTRUCTION: RANDOMLY SELECT ONE STOCK FROM EACH CLUSTER; COMPARE A PORTFOLIO OF 1 SINGLE CLUSTER
```{r}

 #assigns cluster to the stock
stock_clusters <- 
cutree(hc, k = 6) %>%
  as.data.frame() %>% 
  rename(cluster=1) %>% 
  mutate(company_name = rownames(.))

#randomly pick 1 stock in each cluster
randomly_selected_stocks <-  
stock_clusters %>% 
  group_by(cluster) %>% 
  sample_n(size = 1) 

#cluster you want to use as a correlated portfolio
target_cluster <- 4

#randomly pick 6 stocks in your target cluster
individual_cluster <- 
stock_clusters %>% 
  filter(cluster == target_cluster) %>% 
  sample_n(size = 6)


#create a portfolio from stocks selected from each cluster group
portfolio_1 <- 
all_spx_prices %>% 
  filter(security %in% randomly_selected_stocks$company_name) %>% 
  arrange(symbol, date) %>% 
  group_by(symbol) %>% 
  mutate(closing_price = adjusted/lag(adjusted, 1)-1) %>% 
  pivot_wider(id_cols = date, names_from = symbol, values_from=closing_price)

#create another portfolio from stocks selected from a single cluster group
portfolio_2 <- 
all_spx_prices %>% 
  filter(security %in% individual_cluster$company_name) %>% 
  arrange(symbol, date) %>% 
  group_by(symbol) %>% 
  mutate(closing_price = adjusted/lag(adjusted, 1)-1) %>% 
  pivot_wider(id_cols = date, names_from = symbol, values_from=closing_price)

#convert the portfolio data into an XTS object, used by the portfolio analytics package
portfolio_1_xts <- xts(x = portfolio_1[-1], order.by = portfolio_1$date, frequency = 252)
portfolio1_returns <- Return.portfolio(portfolio_1_xts)

#convert the portfolio data into an XTS object, used by the portfolio analytics package
portfolio_2_xts <- xts(x = portfolio_2[-1], order.by = portfolio_2$date, frequency = 252)
portfolio2_returns <- Return.portfolio(portfolio_2_xts)

#calculate the sharpe ratio for both portfolios
SharpeRatio.annualized(R = xts(x = portfolio1_returns, frequency = 252))
SharpeRatio.annualized(R = xts(x = portfolio2_returns, frequency = 252))

#compare the two portfolios - cumulative returns since inception
portfolio1_returns %>% 
  as.data.frame() %>% 
  mutate(date = rownames(.),
         date = ymd(date)) %>%
  rename(DiversifiedPorfolio=1) %>% 
  inner_join(
    portfolio2_returns %>% 
    as.data.frame() %>% 
    mutate(date = rownames(.),
           date = ymd(date)) %>% 
        rename(SingleClusterPorfolio=1)
  ) %>% 
  arrange(date) %>% 
  mutate(DiversifiedPorfolio_Cumulative = cumprod(1+DiversifiedPorfolio)-1,
         SingleClusterPorfolio_Cumulative = cumprod(1+SingleClusterPorfolio)-1) %>% 
  ggplot() +
  geom_hline(aes(yintercept = 0), color = 'black')+
  geom_line(aes(date, SingleClusterPorfolio_Cumulative), color = 'blue') +
  geom_line(aes(date, DiversifiedPorfolio_Cumulative), color = 'orange') +
  labs(title = 'Comparison of Portfolio Returns',
       subtitle = 'Portfolio using hierarchical clustering (orange) vs. Single Cluster (blue)',
       y = 'cumulative return (%)') +
  scale_y_continuous(labels= scales::percent)

# annualized return of the two portfolios
Return.annualized(R = xts(x = portfolio1_returns, frequency = 252))
Return.annualized(R = xts(x = portfolio2_returns, frequency = 252))


```
